
---
title: "Placenta Project"
date: "2022-12-09"
output: html_document
---



Library
```{r}
library("tidyverse")
library("janitor")
library("hrbrthemes")
library("psych")
library("ggrepel")
library("ggplot2")
library("ggpubr")
library("broom")
library("AICcmodavg")
library("ctv") 
library("ape")
library("phytools")
library("phangorn")
library("geiger")
library("nlme")
library("visreg")
library("tidytree")
library("ggthemes")
library("wesanderson")
library("kableExtra")

```

#Data
```{r}
#setting up main df (cancer data)
main_df <- read.csv("main.csv", na.strings = c(" ", "", "NA"))

#Remove rows that contain all NA
main_df <- main_df[rowSums(is.na(main_df)) != ncol(main_df), ]

records_10 <-main_df[main_df$RecordsWithDenominators >= 10,]

#subsetting main df by mammals only 
mammalia <- records_10[records_10$Class == "Mammalia",]

#reading in placenta interdigitation data (invasiveness)
pi <- read.csv("placenta__more_interdigitation.csv", na.strings = c(" ", "", "NA"))

mammals <- read.newick("mammals_species.nwk") # phylogeny data mammalian species


```

#Wrangling Data
```{r}
##following Ashley's renaming of variables for placenta data
pi2 <- pi %>% 
  select(species, placenta_invasiveness, X.4, interdigitation_placenta, X.5)
# Un-reverse score the numeric ordinal variables
#DOUBLE CHECK THIS
pi2$invade = pi2$X.4
pi2$invade[pi2$invade == "1"] <- "4" # used 4 to get these values out of the way for renaming "3's" as "1's"
pi2$invade[pi2$invade == "3"] <- "1"
pi2$invade[pi2$invade == "4"] <- "3"
pi2$digit = pi2$X.5
pi2$digit[pi2$digit == "1"] <- "4" # same here
pi2$digit[pi2$digit == "3"] <- "1"
pi2$digit[pi2$digit == "4"] <- "3"
#select variables with correctly scored numeric ordinal variables
pi3 <- pi2 %>%
  select(species, placenta_invasiveness, invade, interdigitation_placenta, digit)

#renaming variables in neoplasia df for MAMMALS
np2 <- mammalia %>%
  select(Species, NeoplasiaWithDenominators, NeoplasiaPrevalence, Malignant, MalignancyPrevalence, PropMalignant) %>%
  rename(species = "Species") %>%
  rename(tot.neo = "NeoplasiaWithDenominators") %>%
  rename(prev.neo = "NeoplasiaPrevalence") %>%
  rename(tot.mal = "Malignant") %>%
  rename(prev.mal = "MalignancyPrevalence") %>%
  rename(mal.trans = "PropMalignant")

#merge and anti merge mammals with placenta df to combine cancer data with placenta invasiveness
placentacancer <- merge(np2, pi3, by = "species") #mammals is 102
#merge yields 84 mammalian species with complete cancer data and placenta data

antiplacentacancer <- anti_join(np2, pi3, by= "species") #anti_join yields 18 species without full cancer data or placenta data


placentacancer$invade <- as.numeric(placentacancer$invade) # set rates as numeric
placentacancer$digit <- as.numeric(placentacancer$digit)
placentacancer$tot.neo <- as.numeric(placentacancer$tot.neo)
placentacancer$prev.neo <- as.numeric(placentacancer$prev.neo)
placentacancer$tot.mal <- as.numeric(placentacancer$tot.mal)
placentacancer$prev.mal <- as.numeric(placentacancer$prev.mal)
placentacancer$mal.trans <- as.numeric(placentacancer$mal.trans)

#Renaming so graphs look nicer
placentacancer$placenta_invasiveness[placentacancer$placenta_invasiveness == "epi"] <- "Epitheliochorial"
placentacancer$placenta_invasiveness[placentacancer$placenta_invasiveness == "endo"] <- "Endotheliochorial"
placentacancer$placenta_invasiveness[placentacancer$placenta_invasiveness == "hemo"] <- "Hemochorial"

placentacancer2 <- placentacancer %>%
  mutate_all(funs(str_replace(.," ","_"))) # replace all spaces in species names with underscores to be consistent with phylogeny data

```

#Phylogeny Wrangling
```{r}

# Phylogeny - Mammals + Placentas + Cancer (no Life History)
species_placcancer <- placentacancer2 %>%
  dplyr::select("species")

mammals_list <- as.phylo(mammals)

specmammal <- as.data.frame(mammals_list$tip.label)%>%
  rename("species"="mammals_list$tip.label")

specieslist <- merge(specmammal, species_placcancer, by.x = "species", by.y = "species") #merge yields 83 obs. 


specieslist_v <- as.vector(specieslist$species)
specieslist_k <- keep.tip(mammals, specieslist_v)

#writing list of species as csv using phylo 
write.csv(specieslist_v, "specieslist_v.csv", row.names = FALSE, col.names = TRUE)

specieskept <- read.csv("specieslist_v.csv") #list of 83 species 

#merge mammals and placenta cancer data 83 obs. 
mammallist <- merge(specieskept, placentacancer2, by.x = "x", by.y = "species")%>%
  rename("species"=x)

#setting na to 0 and organizing
mammallist1 <- mammallist
mammallist1[is.na(mammallist1)] <- 0

#phylo data in table format
mammal.table <- as_tibble(mammals)
as.phylo(mammals)

```

# Placental Invasion x Cancer ANOVAs
```{r}
# Malignant transformation as a function of placental invasion
#Note: in my code placentacancer is equal to placentcancer1 in ashley's code.
pmtanova <- aov(mal.trans ~ invade, data = placentacancer)
summary(pmtanova) # 22 N/As

# Malignancy prevalence as a function of placental invasion
pmpanova <- aov(prev.mal ~ invade, data = placentacancer)
summary(pmpanova)

# Neoplasia prevalence as a function of placental invasion
pnpanova <- aov(prev.neo ~ invade, data = placentacancer)
summary(pnpanova)


```

# Interdigitation x Cancer ANOVAs
```{r}
# Malignant transformation as a function of interdigitation
imtanova <- aov(mal.trans ~ digit, data = placentacancer)
summary(imtanova)

# Malignancy prevalence as a function of interdigitation
impanova <- aov(prev.mal ~ digit, data = placentacancer)
summary(impanova)

# Neoplasia prevalence as a function of interdigitation
inpanova <- aov(prev.neo ~ digit, data = placentacancer)
summary(inpanova)

```


# Invasion + Interdigitation x Cancer ANOVAs (additive testing)
```{r}
# Malignant transformation as a function of invasion + interdigitation
pimtanova <- aov(mal.trans ~ invade + digit, data = placentacancer)
summary(pimtanova)

# Malignancy prevalence as a function of invasion + interdigitation
pimpanova <- aov(prev.mal ~ invade + digit, data = placentacancer)
summary(pimpanova)

# Neoplasia prevalence as a function of invasion + interdigitation
pinpanova <- aov(prev.neo ~ invade + digit, data = placentacancer)
summary(pinpanova)

```


#Invasion * Interdigitation x Cancer ANOVAs (interaction testing)
```{r}
# Malignant transformation as a function of invasion * interdigitation
pimtanova1 <- aov(mal.trans ~ invade * digit, data = placentacancer)
summary(pimtanova1) #22 N/As

# Malignancy prevalence as a function of invasion * interdigitation
pimpanova1 <- aov(prev.mal ~ invade * digit, data = placentacancer)
summary(pimpanova1)

# Neoplasia prevalence as a function of invasion * interdigitation
pinpanova1 <- aov(prev.neo ~ invade * digit, data = placentacancer)
summary(pinpanova1)

```

# Invasion AICs
```{r}

# Malignant Transformation
mtset <- list(pmtanova, pimtanova, pimtanova1)
mtnames <- c("pmtanova", "pimtanova", "pimtanova1")
aictab(mtset, modnames = mtnames)
        #"pimtanova1",interaction model is the best fit.
par(mfrow=c(2,2))
plot(pimtanova1)
par(mfrow=c(1,1))


# Malignancy Prevalence
mpset <- list(pmpanova, pimpanova, pimpanova1)
mpnames <- c("pmpanova", "pimpanova", "pimpanova1")
aictab(mpset, modnames = mpnames)
      #"pimpanova", additive model is the best fit.
par(mfrow=c(2,2))
plot(pimpanova)
par(mfrow=c(1,1))


# Neoplasia Prevalence
npset <- list(pnpanova, pinpanova, pinpanova1)
npnames <- c("pnpanova", "pinpanova", "pinpanova1")
aictab(npset, modnames = npnames)
      #"pnpanova", solo model is the best fit
par(mfrow=c(2,2))
plot(pnpanova)
par(mfrow=c(1,1))


```


# Interdigitation AICs
```{r}


# Malignant Transformation
mtsetd <- list(imtanova, pimtanova, pimtanova1)
mtnamesd <- c("imtanova", "pimtanova", "pimtanova1")
aictab(mtsetd, modnames = mtnamesd)
      #"imtanova", solo model is the best fit, interactive a close 2nd
par(mfrow=c(2,2))
plot(imtanova)
par(mfrow=c(1,1))
par(mfrow=c(2,2))
plot(pimtanova1)
par(mfrow=c(1,1))


# Malignancy Prevalence
mpsetd <- list(impanova, pimpanova, pimpanova1)
mpnamesd <- c("impanova", "pimpanova", "pimpanova1")
aictab(mpsetd, modnames = mpnamesd)
      #"impanova", solo model is the best fit
par(mfrow=c(2,2))
plot(impanova)
par(mfrow=c(1,1))


# Neoplasia Prevalence
npsetd <- list(inpanova, pinpanova, pinpanova1)
npnamesd <- c("inpanova", "pinpanova", "pinpanova1")
aictab(npsetd, modnames = npnamesd)
      #"inpanova", solo model is the best fit
par(mfrow=c(2,2))
plot(inpanova)
par(mfrow=c(1,1))

```

Color Palette
```{r}
#making color combos for aes
earth <- c("#61bc36", "#365a54", "#444444")
orange <- c("#FF8459", "#CC530A", "#94280A")
blue <- c("#A8D5C0", "#599BA1", "#1E4E64")
```


Placentas & Cancer Graphs
```{r}

# Placental invasion x neoplasia prevalence

placentacancer$placenta_invasiveness <- factor(placentacancer$placenta_invasiveness,levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))
placentacancer %>%
  ggplot(.)+
  geom_violin(aes(x = placenta_invasiveness, y = prev.neo, fill=placenta_invasiveness))+
  ggtitle("Neoplasia Prevalence by Invasiveness")+
  xlab("Placenta Invasiveness")+
  ylab("Neoplasia %")+
  theme_pander()+
  scale_fill_manual(values = orange)+
  theme(legend.position = "none")


# Interdigitation x neoplasia prevalence

placentacancer$interdigitation_placenta <- factor(placentacancer$interdigitation_placenta,levels = c("Villous", "Trabecular", "Labyrinthine"))

placentacancer %>%
  ggplot(.)+
  geom_violin(aes(x= interdigitation_placenta, y= prev.neo, fill=interdigitation_placenta))+
  ggtitle("Neoplasia Prevalence by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Neoplasia %")+
  theme_pander()+
  scale_fill_manual(values = blue)+
  theme(legend.position = "none")


# Placental invasion x malignancy prevalence

placentacancer %>%
  ggplot(.)+
  geom_violin(aes(x = placenta_invasiveness, y = prev.mal, fill=placenta_invasiveness))+
  ggtitle("Malignancy Prevalence by Invasiveness")+
  xlab("Placenta Invasiveness")+
  ylab("Malignant %")+
  theme_pander()+
  scale_fill_manual(values = orange)+
  theme(legend.position = "none")


# Interdigitation x malignancy prevalence

placentacancer %>%
  ggplot(.)+
  geom_violin(aes(x= interdigitation_placenta, y= prev.mal, fill=interdigitation_placenta))+
  ggtitle("Malignancy Prevalence by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Malignant %")+
  theme_pander()+
  scale_fill_manual(values = blue)+
  theme(legend.position = "none")



# Placental invasion x malignant transformation rate

placentacancer %>%
  ggplot(.)+
  geom_violin(aes(x= placenta_invasiveness, y= mal.trans, fill=placenta_invasiveness))+
  ggtitle("Rate of Malignant Transformation by Invasiveness")+
  xlab("Placenta Invasiveness")+
  ylab("Rate of Malignant Transformation")+
  theme_pander(lp= "none")+
  scale_fill_manual(values = orange)+
  theme(legend.position = "none")




# Interdigitation x malignant transformation rate

placentacancer %>%
  ggplot(.)+
  geom_violin(aes(x= interdigitation_placenta, y= mal.trans, fill=interdigitation_placenta))+
  ggtitle("Rate of Malignant Transformation by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Rate of Malignant Transformation")+
  theme_pander(lp="none")+
  scale_fill_manual(values = blue)


```


Placental Invasion x Cancer Correlation Computations
```{r}

# Placental Invasion x Malignant Transformation
invmal1 <- lm(placentacancer$invade ~ placentacancer$mal.trans)
summary(invmal1)

# Placental Invasion x Neoplasia Prevalence
invneo1 <- lm(placentacancer$invade ~ placentacancer$prev.neo)
summary(invneo1)

# Placental Invasion x Malignancy Prevalence
invmalprev1 <- lm(placentacancer$invade ~ placentacancer$prev.mal)
summary(invmalprev1)

```


Interdigitation Correlation x Cancer Computations
```{r}

# Interdigitation x Malignant Transformation
digmal1 <- lm(placentacancer$digit ~ placentacancer$mal.trans)
summary(digmal1)
 
# Interdigitation x Neoplasia Prevalence
digneo1 <- lm(placentacancer$digit ~ placentacancer$prev.neo)
summary(digneo1)

# Interdigitation x Malignancy Prevalence
digmalprev1 <- lm(placentacancer$digit ~ placentacancer$prev.mal)
summary(digmalprev1)

```

#Practicing Linear Regressions
```{r}
#neoplasia prevalence vs adult_weight.g. linear regression
lmNeoWeight = lm(formula = NeoplasiaPrevalence ~ adult_weight.g., data = mammalia)
summary(lmNeoWeight)
#neoplasiaprev vs adult weight plot with logarithmic adult weight! (does this work?)
plot(log(mammalia$adult_weight.g.), mammalia$NeoplasiaPrevalence, pch = 16, col = "red") + abline(lmNeoWeight)

#regression for malignancy prev and adult weight
lmMalWeight = lm(formula = MalignancyPrevalence ~ adult_weight.g., data = mammalia)
summary(lmMalWeight)

plot(log(mammalia$adult_weight.g.), mammalia$MalignancyPrevalence, pch = 16, col = "blue") + abline(lmMalWeight)
#regression plot (with log scale?)


```


PGLS - Placentas + Cancer (Brownian)
```{r}

#line 784

```




